\section{Background}
\label{sec:background}

This project is about lossless compression, so the original file must be recovered exactly after decompression. The main idea used in our work is simple: first reorganize the data so patterns become easier to see, then model the data, and finally encode it into fewer bits.

Our main compression pipeline follows this logic. In the statistical versions, the data can first pass through the Burrows--Wheeler Transform (BWT) \cite{burrows1994} and Move-to-Front (MTF) \cite{bentley1986}. These steps do not compress the file by themselves, but they help group similar values together, which makes the next stage more effective. After that, the compressor uses an adaptive model \cite{cleary1984} and range coding to generate the final compressed output \cite{rissanen1979,schindler1998}.

Not all versions of the project use the same approach. The more balanced versions, especially \texttt{v5} and \texttt{v6}, follow the statistical pipeline described above. In contrast, \texttt{v8} uses a simpler LZ77-style method \cite{ziv1977,colletlz4}. It is much faster, but it usually produces larger files. This comparison is useful because it shows the main trade-off explored in the project: stronger compression usually needs more processing time.

One important idea that came up during the project is the difference between Shannon entropy and Kolmogorov complexity \cite{kolmogorov1965,shannon1948}. Shannon entropy looks at the byte frequencies in the data and tells you how many bits per symbol you need on average. Kolmogorov complexity asks a different question: how short is the shortest program that can produce this data? Most of the time these two measures agree, but they can disagree completely. For example, the output of a pseudorandom number generator looks statistically random---the bytes are uniformly distributed, there are no repeated patterns, and the Shannon entropy is close to the maximum of 8 bits per byte. But the data is entirely determined by a short seed, so a tiny program can reproduce it all. Standard compressors cannot exploit this because they only look for statistical patterns. To compress this kind of data, you need a model that understands the computational process behind it, not just the byte distribution.
