\section{Methodology}
\label{sec:methodology}

\subsection{System Architecture}
The report uses a version-focused methodology. Rather than treating the repository as a single compressor, we compare five representative designs that span the project's design space. \texttt{v9} is the current production version, refining the \texttt{v5} statistical pipeline into the best overall speed/ratio balance. \texttt{v6} extends that pipeline with per-block candidate search for stronger compression. \texttt{v7} and \texttt{v8} explore the speed end of the spectrum, trading ratio for throughput. \texttt{v10} adds PRNG detection on top of \texttt{v6}, demonstrating Kolmogorov compression on pseudorandom data.

\subsection{Algorithm Selection}
The \texttt{v5}/\texttt{v9} line is the primary subject because it best illustrates the repository's statistical compression pipeline: BWT and MTF reshape low-entropy structure before modeling, order-1 adaptive modeling captures local byte dependencies, and range coding/rANS provide the coding stage. \texttt{v6} is included because it improves ratio further through per-block candidate selection over multiple transforms and model orders. \texttt{v7} and \texttt{v8} serve as speed-focused contrasts---\texttt{v7} uses interleaved rANS, while \texttt{v8} drops entropy coding entirely for an LZ77-style matcher. \texttt{v10} is included because it addresses a fundamentally different problem: data that appears incompressible by Shannon entropy but is trivially compressible by Kolmogorov complexity.

\subsection{Optimization Strategy}
Development proceeded iteratively. Earlier versions established the basic order-0 and range-coding path, then introduced BWT preprocessing and faster data structures. The \texttt{v5} line consolidated the strongest ideas into one pipeline, and \texttt{v9} refined it by removing unnecessary Order-2 overhead and fixing the ZRLE guard. \texttt{v7} and \texttt{v8} explored the speed extreme by discarding preprocessing and entropy coding. \texttt{v6} extended the statistical line with per-block candidate search. Finally, \texttt{v10} took a different approach entirely: instead of improving the statistical model, it added a computational model that detects PRNG-generated data and stores only the seed.

\subsection{Benchmarking}
All claims in this report are based on the benchmark corpus used throughout the project. The corpus contains eight files (A--H) totaling 13 MiB, with different characteristics including text, structured binaries, high-entropy data, and pseudorandom PRNG output. The reported metrics are compressed size, compression ratio, bits per symbol, compression time, decompression time, and lossless correctness. The benchmarks compare \texttt{v5}, \texttt{v6}, \texttt{v7}, \texttt{v8}, \texttt{v9}, and \texttt{v10} against gzip, bzip2, xz, and zstd, using the standard release configuration (\texttt{g++}/C++17 with \texttt{-O3 -march=native -flto=auto}).
